---
title: "analysis template"
author: "Zitong Wang"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r Load packages, echo = F, message = F, warning =  F, results = 'hide'}
Packages <- c("forcats", "splines", "dplyr", "tidyr", "tidyverse", "tibble", "ggplot2", "data.table", "RColorBrewer","grid","ggpubr", "gridExtra", "lme4", "nlme")
#install.packages(Packages)  
## load packages
lapply(Packages, library, character.only = TRUE)
```
### Process data


```{r Process data00,  eval = F}
dat.AI<-read.csv('AI_tetall.csv' , header = F)
dat.AU <- read.csv('AU_tetall.csv' , header = F)
dat.Y<- read.csv('Y_tetall.csv' , header = F)
colnames <- c('rat','tetrode', 'distance', 'speed', 'sumcorr') 
names(dat.AI) <- colnames
names(dat.AU) <- colnames
names(dat.Y) <- colnames
pasteit=function(m,collapse=".") {paste(m,collapse=collapse)}
dat.AI$cell.id=apply(cbind(dat.AI$rat, dat.AI$tetrode),1,pasteit)
dat.AU$cell.id=apply(cbind(dat.AU$rat, dat.AU$tetrode),1,pasteit)
dat.Y$cell.id =apply(cbind(dat.Y$rat, dat.Y$tetrode), 1,pasteit)
groups=c("0.Y","1.AU","2.AI")

datnew_Y = dat.Y %>% 
  group_by(cell.id) %>% 
  summarise(rat = rat,
           mspeed = speed,
            mdist = distance,
            #rate = rate,
            #fieldsize = fieldsize,
            corr = sumcorr,
            n=n(),
            group="0.Y")

datnew_AU= dat.AU %>% 
  group_by(cell.id) %>% 
  summarise(rat = rat,
            mspeed = speed,
            mdist = distance,
            #rate = rate,
            #fieldsize = fieldsize,
            corr = sumcorr,
            n=n(),
            group="1.AU")

datnew_AI = dat.AI %>% 
  group_by(cell.id) %>% 
  summarise(rat = rat,
            mspeed = speed,
            mdist = distance,
            #rate = rate,
            #fieldsize = fieldsize,
            corr = sumcorr,
            n=n(),
            group="2.AI")


datnew_all= rbind(datnew_Y,datnew_AU,datnew_AI)
dt <- datnew_all[complete.cases(datnew_all),]
#dt$rat <- substr(dt$cell.id, 1, 6)
dt.tot <- dt 
dt.tot$mspeedm20 <- dt.tot$mspeed - 15 #center speed and distance
dt.tot$mdistm15 <- dt.tot$mdist - 0.5
dt.tot <- as.data.frame(dt.tot)
dt.tot$rat <- factor(dt.tot$rat, 
                     levels = unique(dt.tot$rat[order(dt.tot$group, dt.tot$mdist)]),
                     ordered=TRUE)
```


### Observed correlation along transverse axis organized according to each animal (each dot is a tetrode)


```{r, Observed plot01}
#ggplot(data = dt.tot, aes(y = corr, x = mdist,color = group))+
ggplot(data = dt.tot, aes(y = (corr), x = mdist,color = group))+
  geom_point(aes(group = group), alpha= 0.5)+theme_bw()+
  geom_smooth(aes(group = rat), method = 'lm', se = F)+  
  facet_grid(~rat)+
  ylab('correlation')+
  xlab('Mean Distance')+
  ggpubr::rotate_x_text(angle = 45)+
  theme(strip.text = element_text(size=6))
```

### Models

#### Linear mixed-effect models

  + ModelB: log(rate) ~ group * ns(distance, 2) + ns(speed, 3) + (1 + distance|rat)

```{r, Models00}
df = 2
bknts = range(dt.tot$mdistm15)
p = c(1:(df-1))/df
knts = quantile(dt.tot$mdistm15, p)

df.sp = 3
bknts.sp = range(dt.tot$mspeedm20)
p.sp = c(1:(df.sp-1))/df.sp
knts.sp = quantile(dt.tot$mspeedm20, p.sp)


fm2 <- as.formula(paste('corr ~  group*ns(mdistm15, Boundary.knots = c(', bknts[1],",",bknts[2], "),knots=", knts, ")", "+ns(mspeedm20,Boundary.knots = c(", bknts.sp[1], ",", bknts.sp[2], "),knots=c(",knts.sp[1], ",", knts.sp[2],"))" ))

modB <- lme(fm2,
             random = ~1+mdistm15|rat,
             data = dt.tot,
             control = lmeControl(opt = 'optim'))
fixef(modB)
```


#### Plot model outputs 



```{r, Plot model output01}
groups=c("0.Y","1.AU","2.AI")
Yrat <- unique(dt.tot$rat[dt.tot$group == groups[1]])
AUrat <- unique(dt.tot$rat[dt.tot$group == groups[2]])
AIrat <- unique(dt.tot$rat[dt.tot$group == groups[3]])
ratid <- NULL
ratid <- c(as.character(Yrat), as.character(AUrat), as.character(AIrat))

dist.grid = seq(0.01,1, by = 0.01)-0.5
newdat1 = data.frame(
  #rat=as.factor(rep("160281", 3*length(dist.grid))),
  rat=c(rep(ratid,each = length(dist.grid))),
  mdistm15=rep(dist.grid,18),
  mspeedm20 = rep(c(-1),each = 6*length(dist.grid)),
  group=rep(c("0.Y","1.AU","2.AI"),each = length(dist.grid)))

newdat1$mdist = newdat1$mdistm15 + 0.5
newdat1$fitB <- predict(modB, newdata = newdat1, level = 0) 
newdat1$fitBr <- predict(modB, newdata = newdat1)

Designmat <- model.matrix(eval(eval(modB$call$fixed)[-2]), data = newdat1)

#compute standard error for predictions
predvar <- diag(Designmat %*% modB$varFix %*% t(Designmat))
newdat1$SE <- sqrt(predvar) 

p4 <- ggplot(data = newdat1, aes(x = mdist))+
  geom_line(aes(y = fitB, group = group, linetype = 'Model2 Overall', color = group), size= 1)+
 # geom_line(aes(y = fitBr,group = rat, linetype = 'Model2 Rat-specific', color = group))+
  geom_ribbon(aes(y = fitB, ymin=fitB-1.96*SE,ymax=fitB+1.96*SE, group = group, fill = group),alpha=0.2)+
  ylab('correlation')+
  xlab('Mean Distance')+
  ggtitle('ModelB - Fitted log spike rate vs mean distance')+
  ylim(c(-2, 3))+
  theme_bw()
p4 

```

#### Tests

* Wald tests testing if the shape of the curves is different across age groups. Specifically we test,

  + Average of AU and AI with Y (A - Y)
  
  + AU - AI

  + AU - Y
  
  + AI - Y


```{r, Tests00}
shapetest.dist <- function(mod){
  ## A - Y
  if(class(mod) == 'lmerModLmerTest'|class(mod) == 'lmerMod'){
    smod <- summary(mod)
    fts <- grep(":",rownames(smod[["coefficients"]]))
    coef <- smod[["coefficients"]][fts,1]
    cov<-smod[["vcov"]][fts,fts]
  }else if(class(mod) == 'lme'){
    smod <- summary(mod)
    fts <- grep(":",names(smod[["coefficients"]][['fixed']]))
    
    coef <- smod[["coefficients"]][['fixed']][fts]
    cov<-smod[["varFix"]][fts,fts]
  }
  
  m.lc1 <- matrix(0, nrow = length(coef)/2, ncol = length(coef))
  for(j in 1:(length(coef)/2)){
    m.lc1[j, 2*j-1] <- 1/2
    m.lc1[j, 2*j] <- 1/2
  }
  lc1 <- m.lc1 %*% as.matrix(coef, ncol = 1)
  var.lc1 <- m.lc1 %*% as.matrix(cov) %*% t(m.lc1)
  chisq.lc1 <- t(lc1) %*% solve(var.lc1) %*% lc1
  p.lc1 <- 1-pchisq(as.numeric(chisq.lc1), length(lc1))
  p.lc1 
  ## AU - AI
  m.lc2 <- matrix(0, nrow = length(coef)/2, ncol = length(coef))
  for(j in 1:(length(coef)/2)){
    m.lc2[j, 2*j-1] <- 1
    m.lc2[j, 2*j] <- -1
  }
  lc2 <- m.lc2 %*% as.matrix(coef, ncol = 1)
  var.lc2 <- m.lc2 %*% cov %*% t(m.lc2)
  chisq.lc2 <- t(lc2) %*% solve(var.lc2) %*% lc2
  p.lc2 <- 1-pchisq(as.numeric(chisq.lc2), length(lc2))
  p.lc2 
  
  ## AU - Y
  if(class(mod) == 'lmerModLmerTest' | class(mod) == 'lmerMod'){
    smod <- summary(mod)
    fts <- grep("AU:",rownames(smod[["coefficients"]]))
    coef <- smod[["coefficients"]][fts,1]
    cov<-smod[["vcov"]][fts,fts]
  }else if(class(mod) == 'lme'){
    smod <- summary(mod)
    fts <- grep("AU:",names(smod[["coefficients"]][['fixed']]))
    
    coef <- smod[["coefficients"]][['fixed']][fts]
    cov<-smod[["varFix"]][fts,fts]
  }
  
  lc3 <- as.matrix(coef, ncol = 1)
  var.lc3 <- cov
  chisq.lc3 <- t(lc3) %*% solve(var.lc3) %*% lc3
  p.lc3 <- 1-pchisq(as.numeric(chisq.lc3), length(lc3))
  p.lc3
  
  ## AI - Y
  if(class(mod) == 'lmerModLmerTest'|class(mod) == 'lmerMod'){
    smod <- summary(mod)
    fts <- grep("AI:",rownames(smod[["coefficients"]]))
    coef <- smod[["coefficients"]][fts,1]
    cov<-smod[["vcov"]][fts,fts]
  }else if(class(mod) == 'lme'){
    smod <- summary(mod)
    fts <- grep("AI:",names(smod[["coefficients"]][['fixed']]))
    
    coef <- smod[["coefficients"]][['fixed']][fts]
    cov<-smod[["varFix"]][fts,fts]
  }
  lc4 <- as.matrix(coef, ncol = 1)
  var.lc4 <- cov
  chisq.lc4 <- t(lc4) %*% solve(var.lc4) %*% lc4
  p.lc4 <- 1-pchisq(as.numeric(chisq.lc4), length(lc4))
  p.lc4
  
  temp<-matrix(c(p.lc1, p.lc2, p.lc3, p.lc4), ncol = 1)
  rownames(temp) <- c("A-Y", "AU-AI", "AU-Y", "AI-Y")
  colnames(temp) <- c('Shape test')
  return(temp)
}
```

```{r, Tests01}
tB <- shapetest.dist(modB)
colnames(tB) <- 'ModelB shape test'
round(tB, 2)
```
